---
title: "Data Intake"
---


<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>
```{r global_options, echo = FALSE, include = FALSE}
options(width = 999)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      cache = FALSE, tidy = FALSE, size = "small")
```


## A little background
```{r echo=FALSE}
load("..//2018-1201-parkgrid18.rdata")
```

This is based on Carol's matrix for SY2018-2019.  First some validation numbers to make sure we have the data input in a sensible way.  Tuition levels here are coded as numbers 1 through 20.  These levels correspond to price brackets (we know the levels). It would be better to work from exact numbers.  Based on this data sheet, some numbers to check to see if our abstraction makes sense.  These numbers do not account for vacancies, they do account for faculty/staff tuition remission.  The numbers do not account for inability to collect on contracts.

Total Tuition Explained 
```{r,echo=FALSE}
sum(data$bracketavg) ## Total tuition explained
```
Number of Students
```{r,echo=FALSE}
nrow(data) ## Number of students
```
Enrollment by Grade Distribution
```{r,echo=FALSE}
table(data$grade)  ## student distribution
```

Here's a simple scatter plot with a little jitter around the categorical levels so you can see them stack up.  There's a little transparency that lets you see overlapping dots.  Blue here are your international students.  Red are the tuition levels being removed next year.  The blue lines are the maximum tuition levels.  The jitter moves the values around a bit. You can see the "barbell effect" as you advance in grades.

```{r echo=FALSE}
## plots to look at.
plot(jitter(as.numeric(data$grade)),jitter(data$nlevel,factor=0),cex=2,
 axes=FALSE,ann=FALSE,
 bg=c("#00000055","#0000FF55","#FF000055")[1+(data$nlevel==20)+2*(data$nlevel==1)],
 pch=21,col=NA)
box()
axis(1,at=1:13,labels=c("K",1:12))
axis(2)
title(xlab="Grade Level",ylab="Tuition Level",main="Actual SY18-19, as of October")
tmax<-c(12,14,14,14,14,15,15,16,17,18,19,19,19)
segments(x0=(1:13)-.5,x1=(1:13)+.5,y0=tmax,y1=tmax,lwd=3,col="blue")
abline(h=median(data$nlevel),lwd=2,lty=3)
```

The median tuition is at about the 8th level or $11,000-11,999 (closer to the upper limit at 8.8).  The dotted line reflects this level. 


Actual matrix below for your notes.
```{r echo=FALSE}
table(data$level,data$grade)[c(20,19,15,16,17,18,1:14),]
```

## Messing around with numbers
The optimization problem in front of us is to balance several factors
* Total tuition receivable
* Enrollment numbers by grade
* The difficulty in filling the seats
* An implicit model about financial diversity

For example, consider the following solutions on the next tabs.

### Average Strategy
First, we could just set the tuition to the across the board average.

```{r echo=FALSE}
data2<-data.frame(grade=data$grade,nlevel=mean(data$nlevel),bracketavg=mean(data[,4]))
plot(as.numeric(data2$grade),jitter(data2$nlevel,factor=2),cex=2,ylim=c(1,20),
 axes=FALSE,ann=FALSE,bg="#00000055",pch=21,col=NA)
axis(1,at=1:13,labels=c("K",1:12))
axis(2)
box()
title(xlab="Grade Level",ylab="Tuition Level",main="Average Strategy")
```

Then by construction, the metrics before are similar. 

Total Tuition Explained 
```{r,echo=FALSE}
sum(data2$bracketavg) ## Total tuition explained
```
Number of Students
```{r,echo=FALSE}
nrow(data2) ## Number of students
```
Enrollment by Grade Distribution
```{r,echo=FALSE}
table(data2$grade)  ## student distribution
```

Run through standard concerns.
* There's no structural reason it couldn't be set this way: kindergarten market price is $15,000 which is higher than the set average.  
* Can you find these students?  Probably given that you can fill seats at full price.
* The main downfall is abrogating any interest in economic diversity.

### Barbell Strategy
Assume per grade you give out slots for max tuition rate until you fill the class quota.  Number of remaining seats?


```{r echo=FALSE, quote=""}
## barbell
target<-tapply(data[,4],data$grade,sum)
dom<-data[data$nlevel!=20,]
d3<-data[data$nlevel==20,]
d3<-tapply(d3[,4],d3$grade,sum)
d3[is.na(d3)]<-0
intl<-d3
cap<-c(15640,rep(17100,4),c(18170,18580,19890,20440,21770,22870,22870,22870))

out<-
cbind(target,intl,domestic=target-intl,TuitionCap=cap,AllSeats=table(data$grade),
DomSeats=table(dom$grade),
 IntlSeat=table(data$grade)-table(dom$grade),
 FullPayNeeded=ceiling((target-intl)/cap),
 ZeroPay = table(dom$grade)- ceiling((target-intl)/cap))
round(out)
```

Okay check our work

```{r echo=FALSE}
data3<-data.frame(grade=factor(rep(rownames(out),times=out[,5]),c("K",1:12)),
   price=unlist(apply(out,1,function(x) rep(c(x[4],0,31500),times=c(x[8],x[9],x[7]) ))),
   nlevel=unlist(apply(cbind(tmax,out[,-1]),1,function(x) rep(c(x[1],6,20),times=c(x[8],x[9],x[7]) )))
  )

plot(as.numeric(data3$grade),jitter(data3$nlevel,factor=2),cex=2,
 axes=FALSE,ann=FALSE,bg=c("#00000055","#0000FF55","#FF000055")[1+(data3$nlevel==20)+2*(data3$nlevel==1)],
  pch=21,col=NA)
box()
axis(1,at=c(1:13),labels=c("K",1:12))
axis(2)
title(xlab="Grade Level",ylab="Tuition Level",main="Barbell Strategy")
tmax<-c(12,14,14,14,14,15,15,16,17,18,19,19,19)
segments(x0=(1:13)-.5,x1=(1:13)+.5,y0=tmax,y1=tmax,lwd=3,col="blue")
text(x=1:13,y=6,out[,9],col=2,cex=2)
text(x=1:13,y=9,out[,8],col=1,cex=2)

```


Total Tuition Explained 
```{r,echo=FALSE}
sum(data3$price) ## Total tuition explained
```
Number of Students
```{r,echo=FALSE}
nrow(data3) ## Number of students
```
Enrollment by Grade Distribution
```{r,echo=FALSE}
table(data3$grade)  ## student distribution
```

Numbers on the plot give the exact number of students required (black) and number of free slots open (red).  Can you hit this in practice?  Well it looks like the school is very far off.  Just to recap: you probably couldn't fully fund the school with full pay students. This means the argument that sets full pay and tuition discounted students in opposition is a fallacy.  One might recast this narrative by recognizing that the fact that most of the students pay something (and we can't find this full pay market) is what keeps tuition working.  A different corollary is that the zero pay student number here is roughly the number of vacancies the system could take and break even at the same revenue.

```{r echo=FALSE}
tmp<-cbind(out[,c(8,9)],ActualFull=c(1,0,2,2,7,1,1,2,3,3,5,3,7))
cbind(tmp,PropFull=round(tmp[,3]/tmp[,1],2))
```

Now in practice, there aren't really that many zero-pay students so really, "students paying anything" is what lets us relax the upper bound on the tuition.

### Market Full, Strict Lower Bound strategy
It's unlikely we can drive enrollment assuming we can fill a large order for full pay students.  Further it is unrealistic to fully ignore the contribution of students who pay less than full tuition and more than zero.  On the other extreme, the average model does not meet our need for financial diversity.  Whether the model could be deployed is not settled: one could object that there just aren't enough families in the area to pay the average tuition for all contracts (to be determined with actual numbers, please).  It is likely our final models will try to find the sweet spot where we can deviate from the average model in equal measure to fill seats.  

Suppose now we go back and consdier the average tuition required in a grade if we had the same number of full pay students in each level. 

```{r echo=FALSE}
tmp<-cbind(ActualFull=c(1,0,2,2,7,1,1,2,3,3,5,3,7))
tmp<-cbind(tmp,Intl=out[,7],Seats=as.numeric(out[,5]-tmp-out[,7]),RevRemaining=target-cap*tmp[,1]-intl)
tmp<-cbind(tmp,AvgReq=tmp[,4]/tmp[,3])
round(tmp)
```





