---
title: "Average Strategy"
---


<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>
```{r global_options, echo = FALSE, include = FALSE}
options(width = 999)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      cache = FALSE, tidy = FALSE, size = "small")
```


```{r echo=FALSE}
load("..//2018-1201-parkgrid18.rdata")
tmax<-c(12,14,14,14,14,15,15,16,17,18,19,19,19)

```


### Average Strategy
First, we could just set the tuition to the across the board average.

```{r echo=FALSE}
data2<-data.frame(grade=data$grade,nlevel=mean(data$nlevel),bracketavg=mean(data[,4]))
plot(as.numeric(data2$grade),jitter(data2$nlevel,factor=2),cex=2,ylim=c(1,20),
 axes=FALSE,ann=FALSE,bg="#00000055",pch=21,col=NA)
axis(1,at=1:13,labels=c("K",1:12))
axis(2)
box()
title(xlab="Grade Level",ylab="Tuition Level",main="Average Strategy")
```

Then by construction, the metrics before are similar. 

Total Tuition Explained 
```{r,echo=FALSE}
sum(data2$bracketavg) ## Total tuition explained
```
Number of Students
```{r,echo=FALSE}
nrow(data2) ## Number of students
```
Enrollment by Grade Distribution
```{r,echo=FALSE}
table(data2$grade)  ## student distribution
```

Run through standard concerns.
* There's no structural reason it couldn't be set this way: kindergarten market price is $15,000 which is higher than the set average.  
* Can you find these students?  Probably given that you can fill seats at full price.
* The main downfall is abrogating any interest in economic diversity.
