---
title: "Census Data by Town"
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>
```{r global_options, echo = FALSE, include = FALSE}
options(width = 999)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      cache = FALSE, tidy = FALSE, size = "small")
```

Based on some conversations about market, it may be useful to examine the trends at a township level across Erie county.  I pulled this data from [US Census QuickFacts](https://www.census.gov/quickfacts/fact/map/).  It has the following variables.

```{r echo=FALSE}
cens<-read.csv("..//Census QuickFacts Dec-01-2018.csv",row.names=1,nrow=9)#,colClasses="numeric")
colnames(cens)<-gsub("\\.","",colnames(cens))
colnames(cens)<-gsub("CountyNewYork","",colnames(cens))
colnames(cens)<-gsub("townErie","",colnames(cens))
colnames(cens)<-gsub("cityNewYork","",colnames(cens))
colnames(cens)<-gsub("townNiagara","",colnames(cens))
cens<-as.matrix(cens)
class(cens)<-"numeric"
cens<-t(cens)

as.matrix(colnames(cens))

library(Rtsne)
library(circlize)
set.seed(808)
fit<-Rtsne(scale(cens)[,-4],perplex=1)
Y<-fit$Y
rownames(Y)<-rownames(cens)
KM<-kmeans(Y,center=6)$cluster
KM[KM==6]<-4
KM[KM==3]<-1
```

Roughly the clusters are:
```{r echo=FALSE}
tapply(rownames(Y),KM,c)
```
and their relative compositions are
```{r echo=FALSE}
cx<-t(data.frame(cluster=KM,cens))[,order(KM)]
cx<-t(apply(cx[-1,],1,function(x)tapply(x,cx[1,],median)))
rownames(cx)<-colnames(Z)
cx
```


t-SNE Plots reflect multivariate distance.  I've colored them here by specific factors. Note that this distance omits the use of number of households.  
```{r echo=FALSE}

grad<-colorRamp2(c(-2,0,2),c("purple","grey","orange"))
Z<-scale(cens)
colnames(Z)[8]<-"Language other than English spoken at home"
k<-3
plot(fit$Y,type="n",axes=FALSE,ann=FALSE)
text(fit$Y,label=rownames(Y),xpd=NA,col=grad(Z[,k]))
title(main=colnames(Z)[k])
legend("bottomright",fill=c("orange","purple"),legend=c("High","Low"))

k<-8
plot(fit$Y,type="n",axes=FALSE,ann=FALSE)
text(fit$Y,label=rownames(Y),xpd=NA,col=grad(Z[,k]))
title(main=colnames(Z)[k])
legend("bottomright",fill=c("orange","purple"),legend=c("High","Low"))

k<-9
plot(fit$Y,type="n",axes=FALSE,ann=FALSE)
text(fit$Y,label=rownames(Y),xpd=NA,col=grad(Z[,k]))
title(main=colnames(Z)[k])
legend("bottomright",fill=c("orange","purple"),legend=c("High","Low"))

k<-2
plot(fit$Y,type="n",axes=FALSE,ann=FALSE)
text(fit$Y,label=rownames(Y),xpd=NA,col=grad(Z[,k]))
title(main=colnames(Z)[k])
legend("bottomright",fill=c("orange","purple"),legend=c("High","Low"))

```

That Alden number probably includes the prison population (best guess, households x avg/household - actual population leaves about 1800 unaccounted for: Wende and ECHF make up maybe 1600 max).

There's a variable about foreign born persons I could add.